Mitigating Interest Flooding DDoS Attacks in Named Data Networking

### 2012-12-05 04-design.tex
Semi-finalized fair-queuing description
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@857 df1a90cd-247b-4364-bdaf-287a704eb499
Stats part
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@861 df1a90cd-247b-4364-bdaf-287a704eb499

\subsubsection{Interest limits (physical limits)}
\label{sec:physical limits}

@@

\subsubsection{Interest limits with per-interface fairness}
\label{sec:queuing}

@@

About additional knowledge about interests.

Another unique feature of the NDN architecture is stateful packet forwarding~\cite{adaptive-forwarding} and symmetric flow of Interests and Data packets.  Every time Interests is send out to the Interface, the router will be able to learn various parameters about this Interest.

From the DDoS mitigation point of view, one of the important parameters is to determine from which interface

In order to mitigate a denial of service attack it is vital to know the origin of the attack.  I the origin of the attack is known, then the attack can be easily suppressed and eventually localized.  In our definition of the attack is that

interest of the attackers are not getting satisfied. 
Every router can learn fro m which interfaces attack traffic is coming and do certain actions to limit forwarding of suspicious interest.   In other words, without trying to distinguish legitimate and malicious interests on the basis of the Interest prefix, routers distinguish them based on recent performance.

\subsubsection{Data plane performance tracking}
\label{sec:stats}

@@

\subsubsection{Probabilistic Interest accept}
\label{sec:probabilistic}

@@

\subsubsection{Dynamic adjustments of window-based limits}
\label{sec:dynamic limits}

>>>>

\subsubsection{\textbf{Interest limits (physical limits)}}
\label{sec:physical limits}

@@

\subsubsection{\textbf{Physical limits with per-interface fairness}}
\label{sec:queuing}

@@

While the Interest limit is the key building block to suppress mechanisms of the Interest flooding attack, legitimate Interests need to be somehow prioritized and malicious Interests need to be somehow penalized in order to completely suppress the attack.
That is, instead of processing Interests always based on the first-in-first-serve rule, NDN routers need some basis to treat the incoming Interests differently.
Thus, the primary task in bringing intelligence to the Interest flooding attack mitigation is to proactively distinguish between legitimate and malicious Interests.

To achieve such an Interest differentiation we can leverage another unique feature of the stateful forwarding in NDN---namely, the guaranteed symmetric flow of Interest and Data packets.
In other words, NDN guarantees that a forwarded Interest would always results in receiving a corresponding Data, provided such Data exists and Interest/Data packets were not lost on the way.
Therefore, all Interests that result in Data return can be considered legitimate, while the ones that always timeout should be deemed as malicious.\footnote{Recall that in order to maximize effect of the Interest flooding attack, an adversary expresses a large volume of junk Interests (see Section~\ref{sec:interest flooding}).  Implications of other types of attacks are discussed in Section~\ref{sec:discussion}.}

Unfortunately, the timeout-based differentiation is reactive by nature: one cannot know in advance that an Interest would timeout or bring Data back.
What we can do proactively is to keep an up-to-date statistics of Interests satisfaction ratios (number of forwarded versus number of satisfied Interests), and use this statistics to project outcomes for new Interests.
For example, maintaining independent Interest satisfaction ratio statistics for each incoming interface can be enough to estimate whether an Interest received from a neighbor will bring Data or time out.
Statistics can also be kept at more granular levels (e.g., per outgoing interface, per prefix, etc.), which can further improve the estimation quality.

The devil is always in the details.
From the one hand, such statistics needs to start penalizing adversaries as soon as possible (i.e., negative stats should build up fast).
On the other hand, the positive statistics should not deteriorate too fast (i.e., positive stats should be relatively long-term).
Our preliminary experiments showed that the standard exponentially weighted moving average, performed once a second with $\alpha$ coefficient $e^{-1/30}$, approximately corresponding to a 30-second averaging window, provides a good balance between the two contradictory requirements.

% \subsubsection{\textbf{Data plane performance tracking}}
% \label{sec:stats}

@@

\subsubsection{\textbf{Probabilistic Interest accept}}
\label{sec:probabilistic}

@@

\subsubsection{\textbf{Dynamic adjustments of window-based limits}}
\label{sec:dynamic limits}

### 2012-12-05 algos/fair-queueing.tex
Stats part
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@861 df1a90cd-247b-4364-bdaf-287a704eb499

As we show in Section~\ref{sec:evaluation}, fair queueing provides a partial relief from the Interest flooding attack, allowing legitimate users to successfully fetch Data for 15--20\% of the expressed Interests (compared to 0-10\% without fair queueing).
At the same time, the Physical limits with or without fair queueing allows attackers to send a relatively small volume of Interests in order to significantly impact service for the legitimate users.
Therefore, to successfully solve the problem, we need a more intelligent approach, allowing us to localize the attack traffic as close to the attack origin as possible.

>>>>

As we show in Section~\ref{sec:evaluation}, fair queueing provides a partial relief from the Interest flooding attack, allowing legitimate users to successfully fetch Data for 15--20\% of the expressed Interests (compared to 0-10\% without fair queueing).
At the same time, the Physical limits with or without fair queueing allows attackers to send a relatively small volume of Interests in order to significantly impact service for the legitimate users.
Therefore, to successfully solve the problem, we need a fundamentally different, more intelligent approach, allowing localization of the attack traffic as close as possible to the attack origin.

### 2012-12-05 algos/probabilistic.tex
Correcting algorithms. Far far far from perfect...
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@849 df1a90cd-247b-4364-bdaf-287a704eb499
First part of the design section semi-done
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@855 df1a90cd-247b-4364-bdaf-287a704eb499
Stats part
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@861 df1a90cd-247b-4364-bdaf-287a704eb499

\subsection{Probabilistic Interest accept algorithm}
\label{sec:probabilistic}

@@

\Function{InInterest}{Interest $i$, InFace $if$}

@@

    \For{\textbf{each} OutFace $of$ \textbf{in} ForwardingDecision($i,if$)}
        \State $s \leftarrow if$.GetSatisfactionRatio()
        \If{$s \in [0,1]$} \Comment{A valid stats exists}
            \State Stop processing with probability $P(s)$
        \EndIf
        \If{$of$ is under physical limits}
            \State Forward($i$, $of$)
        \EndIf
    \EndFor
\EndFunction

@@

The definition of the probabilistic Interest accept is very straightforward and seems to be trivial to implement.
However, there are a couple of implementation caveats that complicate matters.
Note that there is a condition (line 6 on Pseudocode~\ref{alg:probabilistic mode}) to check if there is a valid statistics point.
This condition is extremely important, because it first provides a basis to distinguish between known facts (i.e., good or bad satisfaction ratio for the incoming interface) and unknown facts (e.g., the first time an Interests arrives on the interfaces).
Second, it gives an opportunity to recover from a bad history (history of unsatisfied Interests) after malicious Interests are ceased to flow in.
Essentially, this recovery relies on statistics module to perform time-based invalidation of historical data (timely, but not too quickly\footnote{Otherwise, attackers may send short bursts of malicious Interests, successfully avoiding differential Interest treatment}).
As described in Section~\ref{sec:stats}, we are currently experimenting with the exponential statistics degradation, while we also planning to apply different approaches in the future.

>>>>

% \subsection{Probabilistic Interest accept algorithm}
% \label{sec:probabilistic}

@@

\Function{InInterest}{Interest $i$, InInterface $if$}

@@

    \For{\textbf{each} OutInterface $of$ \textbf{in} FwDecision($i,if$)}
        \If{$F_{if} > 1$} \Comment{At least one Interest was forwarded}
            \State $s \leftarrow 1 - \min(0, U_{if} / F_{if})$
            \State Drop interest with probability $P(s)$
        \EndIf

        \State{forward the Interest, subjecting to physical limits}
    \EndFor
\EndFunction

@@

% The definition of the probabilistic Interest accept is very straightforward and seems to be trivial to implement.
% However, there are a couple of implementation caveats that complicate matters.
